
# Reproducible Research Assignment Project 1



### Loading and preprocessing the data


Code for reading the activity data from csv file:


```{r read}
activity <- read.csv('activity.csv')
```


Upon inspection, the interval start minutes jump from x55 minutes to x00. Apparently, hours are erroneously counted as having 100 minutes.The following code fixes this flaw:


```{r fix minutes}
activity$interval <- (activity$interval %/% 100) * 60 + activity$interval %% 100
```


### What is mean total number of steps taken per day?


Code for calculating the total number of steps per day and displaying a histogram:


```{r totalsteps}
perday <- tapply(activity$steps,activity$date,sum)
hist(perday,main = 'Histogram of steps per day',xlab = 'steps per day')

```


Code for calculating the mean and median number of steps per day:


```{r meansteps}
mean <- mean(perday,na.rm=TRUE)
median <- median(perday,na.rm=TRUE)
```

Mean steps per day: `r mean`

Median steps per day: `r median`



### What is the average daily activity pattern?


Code for calculating the average of each 5 minute interval and making a time series plot of it:


```{r time series}
meanperinterval <- aggregate(activity$steps,list(activity$interval),mean,na.rm = TRUE)
colnames(meanperinterval) <- c("interval","steps")
maxint <- which.max(meanperinterval$steps)
library(ggplot2)
ggplot(meanperinterval,aes(y=steps,x=interval)) +
  geom_line() +
  geom_point(x=meanperinterval[maxint,1],y=meanperinterval[maxint,2],color="red")
```

The `r maxint`th 5-minute interval, starting at `r meanperinterval[maxint,1]` minutes and marked red in the graph, contains the maximum number of steps


### Imputing missing values

```{r missing values}
missing <- is.na(activity$steps)
numbermissing <- sum(missing)
imputed <- activity  #copy original activity dataframe
# replace missing values with mean of that interval taken from meanperinterval data frame
imputed[which(missing),'steps'] <- round(meanperinterval[(which(missing) - 1) %% 288 + 1,'steps']) #use modulo to get daily interval index from total index, round to get integer number of steps

imputedperday <- tapply(imputed$steps,imputed$date,sum)
hist(imputedperday,main = 'Histogram of steps per day',xlab = 'steps per day')
imputedmean <- mean(imputedperday,na.rm=TRUE)
imputedmedian <- median(imputedperday,na.rm=TRUE)
```

Mean steps per day in imputed data: `r imputedmean` as opposed to `r mean` in original data.

Median steps per day in imputed data: `r imputedmedian` as opposed to `r median` in original data.

The impact of imputing on the mean is small and due to the rounding to integer values. 
The impact on the median is small, too.

### Are there differences in activity patterns between weekdays and weekends?

Code to calculate interval averages for weekends and weekdays separately:

```{r weekends weekdays}
activity$day <- factor(weekdays(as.Date(activity$date,"%Y-%m-%d")) %in% c('Saturday','Sunday'), levels=c(TRUE, FALSE), labels=c('weekend', 'weekday') )
meanperintday <- aggregate(activity$steps,list(activity$day,activity$interval), mean, na.rm=TRUE)
colnames(meanperintday) <- c('day','interval','steps')

ggplot(meanperintday,aes(y=steps,x=interval,group=day)) +
  geom_line() +
  facet_wrap(~ day)
```